\chapter{Conclusion \& Future Work}

\section{Conclusion}

This thesis presented \textbf{AniUnFlow}, a structure-first unsupervised optical-flow framework for 2D animation. The central idea is that animation correspondence should not be treated as a purely dense matching problem. Instead, AniUnFlow combines SAM-guided object structure, temporal slot memory, dense local refinement, propagated structural agreement, and boundary-aware residual correction so that motion estimation remains both coherent and expressive in stylized scenes.

The most important outcome of this work is not a single scalar benchmark value, but a clearer design principle for animation flow estimation. Object structure is valuable because it organizes the motion hypothesis in flat, ambiguous regions. Dense refinement is valuable because it restores local articulation and geometric detail that an object prior alone cannot capture. Propagation memory is valuable because it tells the framework when nearby temporal structure is still trustworthy after motion has already been estimated. Together, these components produce a framework that is easier to interpret, easier to visualize, and more aligned with the practical failure modes of anime-style motion. The added interpretability analysis further suggests that AniUnFlow is not only a flow estimator, but also a framework whose internal reasoning can be inspected through structurally meaningful intermediate signals.

Another contribution of the thesis is presentational. By rewriting the work around one shared framework name and adding more schematic and qualitative figures, the paper now reads less like a collection of disconnected experiments and more like a coherent research story. This matters because the development of AniUnFlow was inherently iterative: the architecture improved not only by adding modules, but by learning how structure, memory, and dense refinement should interact.

\section{Limitations and Future Work}

AniUnFlow still leaves several open directions for future work. The most important limitations are not accidental implementation gaps; they reveal which parts of animation motion remain fundamentally hard even under a structure-first formulation.

\paragraph{Very large displacement remains challenging.}
The current framework improves coherence and local detail recovery, but it is still most reliable when a conservative structural prior can be refined rather than completely re-discovered. In clips with extremely large motion, long jumps, or abrupt viewpoint-like changes, the local dense refinement stage may simply not have a strong enough search range to recover the correct correspondence once the initial hypothesis is too far away. This is precisely why the latest large-motion extension is an important continuation of the thesis rather than a side experiment. The next logical step is to combine the existing object-memory design with broader global search, stronger visibility reasoning, and non-rigid slot motion that can better handle large deformations before local refinement begins.

\paragraph{Stylized animation effects remain under-modeled.}
AniUnFlow is designed for ordinary correspondence between successive drawings, but animation often contains frames that intentionally violate ordinary motion assumptions. Smear frames, impact frames, motion streaks, and highly exaggerated transitional poses are not merely ``hard examples''; they can represent deliberate drawing conventions in which the intermediate image is not meant to preserve one-to-one geometric structure at all. In these situations, even a well-regularized flow model may produce visually plausible but semantically questionable correspondences. A stronger future system would therefore need animation-aware effect modeling, for example by detecting special-effect frames explicitly, routing them through a different matching regime, or learning when a dense optical-flow interpretation should be softened in favor of event-aware temporal reasoning.

\paragraph{Propagation memory is still only an indirect structural cue.}
The propagated agreement map introduced in AniUnFlow is useful because it tells the refiner where nearby temporal structure is likely to remain trustworthy. However, it is still a heuristic signal derived from masks and current motion estimates rather than a fully learned uncertainty model. Future work could replace this hand-designed agreement pathway with a learned reliability predictor that explicitly estimates when structural propagation should be trusted, suppressed, or fused with dense evidence. This would make the framework more adaptive in clips containing partial occlusion, region splitting, or inconsistent segment geometry.

\paragraph{Interpretability should move from readable signals to causal explanation.}
One practical strength of AniUnFlow is that it exposes interpretable intermediate states, such as structural partitions, disagreement maps, and residual activity. Yet the current thesis still studies these signals mainly through descriptive visualization. A stronger explainability agenda would go further by perturbing individual masks, ablating propagation-memory guidance at inference time, or tracing how specific slots influence the final flow in selected regions. Such experiments would help distinguish whether a given component is merely correlated with a prediction or actually causally responsible for it.

\paragraph{Downstream evaluation should be expanded into a fuller application study.}
One of the clearest lessons of this thesis is that optical flow for animation should not be judged only as a benchmark output. In practice, flow is valuable because it supports other operations: frame interpolation, inbetweening assistance, temporally consistent editing, mask transfer, color propagation, contour stabilization, and stylization transfer. The present thesis already includes several downstream demonstrations, but they are still lightweight case studies rather than a full application benchmark. A stronger future evaluation should therefore report these tasks more systematically: with multiple clips, clearer task protocols, and, where possible, both perceptual and task-specific metrics. This is particularly important for video interpolation and temporally consistent editing, since these are among the most immediate real-world uses of a stable animation-oriented motion prior.

\paragraph{A broader application direction.}
Taken together, these limitations suggest that the next stage of AniUnFlow should not only seek better endpoint error, but should also move closer to production-oriented animation tools. A more complete evaluation suite would connect flow quality to downstream usefulness more directly, for example by measuring how well the framework supports intermediate-frame synthesis, persistent object-aware editing, effect propagation, and temporally stable stylization under increasingly difficult artistic motion. In that sense, the most meaningful extension of AniUnFlow is not simply ``a bigger model,'' but a more application-aware framework that understands when structure is reliable, when stylization conventions break ordinary correspondence, and when flow should serve as a controllable primitive for higher-level animation assistance.

In summary, AniUnFlow provides a practical foundation for animation-oriented optical flow by showing that structure and dense correspondence should cooperate rather than compete. The framework already supports richer visualization and more readable analysis than the earlier draft of the thesis, and it opens a clear path toward future work on larger motion, stronger temporal reasoning, and production-oriented animation tools.
