\chapter{Introduction}

The animation industry---especially 2D anime production---relies heavily on manual, frame-by-frame motion refinement, making automation increasingly important for scaling modern pipelines across film, television, and digital media. Optical flow plays a central role in this process by providing dense, pixel-level motion correspondences between frames. Such correspondences enable a range of core tasks in anime-style content creation, including high-quality frame interpolation~\cite{animeinterp,superSloMo,softmaxSplat}, inbetweening for line-art sequences~\cite{linedrawingFlow}, motion-consistent color propagation in cel-style painting~\cite{toonTracking}, and improving temporal coherence in stylized rendering and video-to-anime systems. Recent tools such as AnimeInterp~\cite{animeinterp} highlight that reliable optical flow is often the backbone for preserving the distinctive temporal structure of 2D anime, underscoring the need for flow estimators that are tailored to animation-specific visual properties.

Optical flow has been extensively studied in computer vision. Influential supervised methods such as PWC-Net~\cite{pwcnet}, RAFT~\cite{raft}, GMA~\cite{gma}, and GMFlow~\cite{gmflow} achieve strong performance on natural-scene benchmarks including Sintel~\cite{sintel} and FlyingThings3D~\cite{ft3d}. Despite their success in natural videos, these models often generalize poorly to 2D animation. Anime differs fundamentally from natural imagery: it contains large flat-color regions with minimal texture, explicit contour lines that define sharp motion boundaries, exaggerated non-rigid motion and stylized dynamics, and multi-layered compositions with discrete occlusion boundaries. The recently introduced AnimeRun dataset~\cite{animerun} systematically demonstrates that state-of-the-art flow models trained for natural scenes can exhibit severe drift in textureless areas, unreliable handling of large stylized displacements, and limited temporal stability when applied to animation sequences.

A key obstacle to advancing optical flow for animation is the difficulty of obtaining high-quality ground-truth motion. Producing dense flow labels for stylized content typically requires specialized rendering pipelines, careful control over occlusions and camera motion, and extensive curation, which makes large-scale labeled data costly and difficult to diversify. This motivates a shift toward \emph{unsupervised} or \emph{self-supervised} optical flow methods, which learn correspondences from raw videos and can, in principle, scale to long and diverse animation clips without requiring manual annotation.

However, existing unsupervised optical flow methods face notable limitations on animation. First, they rely heavily on photometric consistency, which is frequently violated by flat cartoon regions, stylized shading, and artist-driven appearance changes. Second, most approaches are fundamentally two-frame, lacking the temporal context required to resolve occlusions, disambiguate repeated patterns, and stabilize motion under exaggerated deformations. Third, they typically do not maintain long-range cost representations or memory structures, which makes them susceptible to drift and identity inconsistency in long shots with persistent character trajectories or camera motion. These challenges highlight the need for an architecture that explicitly incorporates multi-frame temporal reasoning and long-range correspondence aggregation for stylized animation.

To address these issues, we introduce \textbf{AniFlowFormer-T}, an unsupervised multi-frame temporal cost-memory Transformer designed specifically for optical flow in 2D animation. AniFlowFormer-T processes a temporal window in a single pass, constructs temporal cost volumes beyond pairwise matching, accumulates long-range correspondence cues in a latent memory, and fuses them through a global temporal regressor. By combining local matching evidence with sequence-level context, the proposed model improves correspondence in flat-color regions, enhances robustness under large stylized motions and occlusions, and promotes temporally coherent flow across extended sequences.

\textbf{Our main contributions are summarized as follows:}
\begin{itemize}
\item We propose a multi-frame temporal cost-memory Transformer for optical flow in 2D animation, enabling long-range spatio-temporal reasoning beyond conventional two-frame designs.
\item We develop an unsupervised training framework that integrates animation-aware photometric reconstruction, occlusion-aware masking, contour-preserving smoothness, and temporal consistency constraints.
\item We evaluate on AnimeRun~\cite{animerun} and show consistent improvements over strong baselines, with especially large gains in texture-sparse regions and under large motion, demonstrating the benefit of explicit temporal memory and global temporal regression.
\end{itemize}
